Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-11 paper/main.tex
minor modifications

\iffull{We only show a small selections of the programs that we can
  handle automatically here.}{} \iffull{In \pref{app:cat}}{In the
  extended version~\cite{anon_extended}} is a list of more than $30$
classes of challenging programs that we can automatically analyze.
\pref{sec:exper} contains a more detailed comparison with other
tools\iffull{for automatic bound derivation.}{.}

@@

The rules of our derivation system must serve two purposes.  They must
\emph{attach} potential to certain program variable intervals and use
this potential, when it is allowed, to \emph{pay} for resource consuming
operations.  These two purposes are easily illustrated on the {\sc Q:Skip}
rule.  This rule reuses its precondition as postcondition.  This is
explained by two facts: first, no resource is consumed by the skip
operation, thus no potential needs to be taken out of the precondition
to pay for the evaluation; second, the program state is not changed by
the execution of a skip statement, thus all the potential available
before the execution of the skip statement is still available after.
These two observations would also justify a rule with the side-condition
$Q' \le Q$, because some potential is always safely lost.  For the sake
of orthogonality, it is not how we present the rule and this form of
weakening is in fact expressed as the stand-alone rule {\sc Q:Weak}.

@@

On the other hand, it might be the case that even with annotations, a
resource bound cannot be found automatically.  With existing tools, this
situation requires a full rewrite of the program.  However, in our
framework, it is possible to write a manual proof and combine it with
bounds derived by the automatic analysis.  This interaction is made
formal using our Quantitative Hoare Logic.  This resource aware program
logic is an extension of the one presented in~\cite{veristack14} and
was originally designed to prove the soundness of our automatic
analysis.  To make it suitable for this end, we made it independent
of the resource used (e.g. stack usage, heap usage, clock ticks) and
implemented a smoother handling of the auxiliary state through a shallow
embedding in Coq.  Despite being first a theoritical tool, the logic can
be used directly as in~\cite{veristack14}.  Its judgements are inspired
by typical Hoare logics and have the shape
$$
\Delta \vdash \htriple P S Q
$$
where $\Delta$ is a set of hypothesis on program functions, and $P, Q$
are quantitative assertions, i.e. functions from the program states
to $\Qplusz \cup \{\infty\}$.  The semantics of triples and the rules
of the logic are described in further details in the extended version of
this paper~\cite{anon_extended}.  Because our automated tool produces
proofs of the resource bounds in the Quantitative Hoare Logic, it can
be used as an assistant when deriving a complex bound by hand in the
logic.  The interaction is guaranteed to be sound by the soundness
theorem of our Quantitative Hoare Logic.

>>>>

\iffull{We only show a small selections of the programs that we can
  handle automatically here.}{} \iffull{In \pref{app:cat}}{In the
  extended version~\cite{anon_extended}} is a list of more than $30$
classes of challenging programs that we can automatically analyze.
\pref{sec:exper} contains a more detailed comparison with other
tools\iffull{ for automatic bound derivation.}{.}

@@

The rules of our derivation system must serve two purposes.  They must
\emph{attach} potential to certain program variable intervals and use
this potential, when it is allowed, to \emph{pay} for resource consuming
operations.  These two purposes are illustrated on the {\sc Q:Skip}
rule.  This rule reuses its precondition as postcondition, it is
explained by two facts: first, no resource is consumed by the skip
operation, thus no potential has to be used to pay for the evaluation;
second, the program state is not changed by
the execution of a skip statement, thus all the potential available
before the execution of the skip statement is still available after.
These two observations would also justify a rule with the side-condition
$Q \ge Q'$, because some potential is always safely lost.  For the sake
of orthogonality, it is not how we present the rule and this form of
weakening is in fact expressed as the stand-alone rule {\sc Q:Weak}
described later.

@@

On the other hand, it might be the case that even with annotations, a
resource bound cannot be found automatically.  With existing tools, this
situation requires a full rewrite of the program.  However, in our
framework, it is possible to write a manual proof and combine it with
bounds derived by the automatic analysis.  This interaction is made
formal using our Quantitative Hoare Logic.  This resource aware program
logic is an extension of the one presented in~\cite{veristack14} and
was originally designed to prove the soundness of our automatic
analysis.  To make it suitable for this end, we made it independent
of the resource used (e.g. stack usage, heap usage, clock ticks) and
implemented a smoother handling of the auxiliary state through a shallow
embedding in Coq.  Despite being first a theoretical tool, the logic can
be used directly as in~\cite{veristack14}.  Its judgements are inspired
by typical Hoare logics and have the shape
$$
\Delta \vdash \htriple P S Q
$$
where $\Delta$ is a set of hypothesis on program functions, and $P, Q$
are quantitative assertions, i.e. functions from the program states
to $\Qplusz \cup \{\infty\}$.  The semantics of triples and the rules
of the logic are described in further details in the extended version of
this paper~\cite{anon_extended}.  Because our automated tool produces
proofs of the resource bounds in the Quantitative Hoare Logic, it can
be used as an assistant when deriving a complex bound by hand in the
logic.  The interaction is guaranteed to be sound by the soundness
theorem of our Quantitative Hoare Logic.
